---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
Let's download and load the data
Create a directory NHTS in the project explorer window
Execute the code below
Check your console for an input

```{r}
library(summarizeNHTS)
download_nhts_data("2017", exdir="NHTS")
dataset <- read_data("2017", csv_path="NHTS")

```

On linux, files need to be rename manually since the script doesn't work. The messages can be found in the console upon running the above code.

Appropriate files need to be renamed as follows:
hhpub -> household
perpub -> person
vehpub -> vehicles

Also note the weight-files in the folder. They are used in Estimation. The package handles the calculations.


```{r}
dataset <- read_data("2017", csv_path="NHTS")

```

Now we have the data loaded in the variable dataset
The variable can be explored in the Environment tab on the right side. Note the 4 tables in dataset$data

Here is a sample query
You should see a plot if it is sucessfully executed

```{r}
dataset <- read_data("2017", csv_path="NHTS")
statistic <- summarize_data(
    data = dataset,
    agg = "household_count",
    by = c("HHSIZE","HHVEHCNT")
)
make_chart(statistic)
```

Let's explore the variable

```{r}
summary(dataset$data)
```

Note the 4 tables corresponding to the NHTS data.
+ trip
+ person
+ household
+ vehicle

Accessing Raw Data:

```{r}
dataset$data$vehicle
```

```{r}
# By position
dataset$data$vehicle[, c(1, 3)]

# By name (single variable)
dataset$data$vehicle$ANNMILES

# By name
dataset$data$vehicle[, list(HOUSEID, ANNMILES)]
```

Selecting Specific Rows

```{r}
# By row numbers (first 5 rows)
dataset$data$vehicle[1:5, ]

# By condition
dataset$data$vehicle[VEHTYPE == "01", ]

# By condition (multiple values)
dataset$data$vehicle[VEHTYPE %in% c("01","02"), ]
```

Print MAKE and MODEL of vehicles
Use NHTS Code book and User guide

```{r}

```

Now let's study summarizeData function
```{r}
summarize_data(
  data = dataset,
  agg = "household_count"
  )
```

Group By:

```{r}
summarize_data(
  data = dataset,
  agg = "household_count",
  by = "HOMEOWN"
)
```

Group by with multiple columns

```{r}
summarize_data(
  data = dataset,
  agg = "household_count",
  by = c("HH_RACE","HOMEOWN")
)
```
```{r}
temp <- summarize_data(
  data = dataset,
  agg = "household_count",
  by = c("HH_RACE","HOMEOWN")
)
make_table(temp)
```

Generating Frequencies

```{r}
# Proportion of persons by WORKER, worker status
temp <- summarize_data(
  data = dataset,
  agg = "person_count",
  by = "WORKER",
  prop = TRUE
)
make_table(temp)
```

Numeric Aggregates

```{r}
temp<-summarize_data(
  data = dataset,
  agg = "avg",
  agg_var = "TRPMILES"
)
make_table(temp)
```

Excercise:

Generate Median TripMiles by Homeownership

```{r}

```

```{r}
summarize_data(
  data = dataset,
  agg = "person_trip_rate",
  by = "WORKER"
)
```

Subsetting Data

```{r}
#Distribution of Leisure Travel
summarize_data(
  data = dataset,
  agg = "trip_count",
  by = "TRAVDAY",
  prop = TRUE,
  subset = "WHYTRP90 %in% c('07','08','10')"
)
```

```{r}
?summarize_data
```